### 2016-08-25 docs/article.Rmd
articolo

---
title: "R package qtdshift: A functional programming approach to shifted probability ditributions"
author: "Veronica"
date: "August 25, 2016"
output: html_document
---

## Introduction

This paper aims to be an implementation of the methodology used by Andrea Spano article, "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach. I was inspired by his work to implement the same methodology to shifted probability distributions.

## Shifted probability function in R

Why shifted distributions? Shift is distribution trasformation used daily in many statistical fields. R does not provide a specific package for shifted distribution models, but it provides package `stats` which contains a wide set of functions for standard distribution models. I refer to: probability, density, quantile and random number generator functions, that share a stable naming convention both for the names of the functions and in the first argument of these functions. In particular:

* probability distribution functions: `p<dist>(q)` with `q` vector of quantiles;
* density functions: `d<dist>(x)` with `x` vector of quantiles;
* quantile functions: `q<dist>(p)` with `p` vector of probabilities;
* random number generation: `r<dist>(n)` with `n` number of observations.

where `<dist>` indicates the conventional name of distribution family.

Let us see an example.  
Considering the Normal distribution (`norm`), we have: `pnorm(q)`, `dnorm(x)`, `qnorm(p)` and `rnorm(n)`.

So, we use to write:

```{r dnorm}
dnorm(x = 6:10, mean = 8, sd = 1)
```

to get density values at 6:10 from a normal distribution with parameters `mean = 8` and `sd = 1`.

Going back to shifted distribution, we could easily write a shifted version for `dnorm()`, as:

```{r}
sdnorm <- function(x, mean = 0, sd = 1, shift = 0, ...){
            x_shifted <- x - shift
            density <- dnorm(x = x_shifted, mean = mean, sd = sd, ...)
            return(density)
}
```

So, we write:

```{r}
sdnorm(x = 8:12, mean = 10, sd = 1, shift = 1)
```

to get density values at 6:10 from a normal distribution with parameters `mean = 8` and `sd = 1`, shifted of 1 unit.

At first view, this approach seems easy and cool but thinking about it I realized that I would need to write a different function for each probability distribution I would like to work with. So I decided to use functional programming approach to develop four general functions which compute shifted distribution models for any distribution chosen:

* `pshift()` for shifted probability distributions;
* `dshift()` for shifted density functions;
* `qshift()` for shifted quantile functions;
* `rshift()` for shifted random numbers generator functions.

These functions are included in `qtdshift` package available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).

They share the same logic: they take as input a probability distribution as a character string and return the equivalent shifted distribution as a function object.

```{r, eval = FALSE}
devtools::install_github("veronicagiro/qshift", build=FALSE)
require(qtdshift)
spnorm <- pshift("norm")
sdnorm <- dshift("norm")
sqnorm <- qshift("norm")
srnorm <- rshift("norm")
```

<!--
esempi
-->