### 2016-08-26 docs/article.Rmd
articolo ok

---
title: 'R package qtdshift: A functional programming approach to shifted probability
  ditributions'
author: "Veronica"
date: "August 25, 2016"
output: html_document
---

## Introduction

This paper aims to be an implementation of the methodology used by Andrea Spanò article, "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach. I was inspired by his work to implement the same methodology to shifted probability distributions.

## Shifted probability functions in R

Why shifted distributions? Shift is a distribution trasformation used daily in many statistical fields. R does not provide a specific package for shifted distribution models, but it provides package `stats` which contains a wide set of functions for standard distribution models. I refer to: probability, density, quantile and random number generator functions, that share a stable naming convention both for the names of the functions and in the first argument of these functions. In particular:

* probability distribution functions: `p<dist>(q)` with `q` vector of quantiles;
* density functions: `d<dist>(x)` with `x` vector of quantiles;
* quantile functions: `q<dist>(p)` with `p` vector of probabilities;
* random number generation: `r<dist>(n)` with `n` number of observations.

where `<dist>` indicates the conventional name of distribution family.

Let us see an example.  
Considering the Normal distribution (`norm`), we have: `pnorm(q)`, `dnorm(x)`, `qnorm(p)` and `rnorm(n)`.

So, we use to write:

```{r dnorm}
dnorm(x = 6:10, mean = 8, sd = 1)
```

to get density values at 6:10 from a Normal distribution with parameters `mean = 8` and `sd = 1`.

Going back to shifted distribution, we could easily write a shifted version for `dnorm()`, as:

```{r sdnorm_definition}
shift_dnorm <- function(x, mean = 0, sd = 1, shift = 0, ...){
            x_shifted <- x - shift
            density <- dnorm(x = x_shifted, mean = mean, sd = sd, ...)
            return(density)
}
```

So we replicate the previous example but shifting the Normal distribution of 1 unit:

```{r sdnorm_example_1}
shift_dnorm(x = 6:10, mean = 8, sd = 1, shift = 1)
```

At first view, this approach seems easy and cool but thinking about it I realized that I would need to write a different function for each probability distribution I would like to work with. So I decided to use functional programming approach to develop four general functions which compute shifted distribution models for any distribution chosen:

* `dshift()` for shifted density functions;
* `pshift()` for shifted probability distributions;
* `qshift()` for shifted quantile functions;
* `rshift()` for shifted random numbers generator functions.

These functions are included in `qtdshift` package available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).

```{r github, eval = FALSE}
devtools::install_github("veronicagiro/qshift", build=FALSE)
```

They share the same logic: they take as input a probability distribution name, as a character string, and return the equivalent shifted distribution as a function object.

```{r qtdshift, message=FALSE}
require(qtdshift)
sdnorm <- dshift("norm") # density function
spnorm <- pshift("norm") # probability function
sqnorm <- qshift("norm") # quantile function
srnorm <- rshift("norm") # random generation function
```

We see in detail the computational process.   
Let us consider `dshift()`. As we seen, it takes as input the distribution name (`norm`) and proceeds as follows:

* gets the corresponding function `dnorm()`;
* uses `dnorm()` to create a function, say `density()` that computes shifted probability for normal distributions;
* modify the formals of `density()` so that it has the same formals as `dnorm()` plus `shift`, corresponding to the shift value;
* returns `density()`.

So, the function `sdnorm()` has the same formals of the function `dnorm()`, plus the parameter `shift` which indicates the shift value to be applied to distribution; by default it is set as 0.

```{r args}
args(sdnorm)
args(dnorm)
```

Package `qtdshift` is based on two key concepts being part of the functional nature of the R programming language:

* the environment of a function can be used as a placeholder for other objects;
* function formals can be manipulated.

Theoretical references in chapters [Environments](http://www.quantide.com/website/ramarro-chapter-03/) and [Functions](http://www.quantide.com/website/ramarro-chapter-04/) of Andrea Spano weeb Book [Ramarro](http://www.quantide.com/website/ramarro-r-for-developers/), available on Quantide web site.

### Examples

Let us consider the Normal distribution and the previously defined functions `sdnorm()`, `spnorm()`, `sqnorm()` and `srnorm()`.

* Density function, `sdnorm()`.  
    This function has a double use:

    * generate probability values from a non-shifted normal distribution by leaving parameter `shift` set to its default (`shift = 0`)
    * generate probability values from any shifted normal distribution by setting `shift` value

    ```{r sdnorm_example}
x <- seq(4, 14, len = 100)
not_shifted <- sdnorm(x = x, mean = 8, sd = 1.5)
positive_shifted <- sdnorm(x = x, mean = 8, sd = 1.5, shift = 1)
negative_shifted <- sdnorm(x = x, mean = 8, sd = 1.5, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r sdnorm_plot, message=FALSE}
require(ggplot2)
df <- data.frame(x=x, 
                 sdnorm_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sdnorm_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Density plot for shifted and non-shifted normal distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Probability function, `spnorm()`:

    ```{r spnorm_example}
not_shifted <- spnorm(q = x, mean = 8, sd = 1.5)
positive_shifted <- spnorm(q = x, mean = 8, sd = 1.5, shift = 1)
negative_shifted <- spnorm(q = x, mean = 8, sd = 1.5, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r spnorm_plot}
df <- data.frame(x=x, 
                 spnorm_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=spnorm_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Quantile") + ylab("Probability") +
  ggtitle("Probability plot for shifted and non-shifted normal distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Quantile function, `sqnorm()`:

    ```{r sqnorm_example}
    x <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
not_shifted <- sqnorm(p = x, mean = 8, sd = 1.5)
positive_shifted <- sqnorm(p = x, mean = 8, sd = 1.5, shift = 1)
negative_shifted <- sqnorm(p = x, mean = 8, sd = 1.5, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r sqnorm_plot}
df <- data.frame(x=x, 
                 sqnorm_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sqnorm_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  ylab("Quantile") + xlab("Probability") +
  ggtitle("Quantile plot for shifted and non-shifted normal distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Random generator function, `srnorm()`:

    ```{r srnorm_example}
    x <- 10000
not_shifted <- srnorm(n = x, mean = 8, sd = 1.5, set_seed = T)
positive_shifted <- srnorm(n = x, mean = 8, sd = 1.5, shift = 3, set_seed = T)
negative_shifted <- srnorm(n = x, mean = 8, sd = 1.5, shift = -3, set_seed = T)
    ```
    We can visualize these results by plotting them:

    ```{r srnorm_plot}
df <- data.frame(srnorm_type = factor(c(rep("not_shifted", times=x),
                    rep("positive_shifted", times=x),
                    rep("negative_shifted", times=x)),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- pl <- ggplot(data = df, mapping = aes(x=value, group=srnorm_type, fill=srnorm_type)) +
  geom_density(alpha=0.6) +
    scale_fill_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  ggtitle("Density plot for shifted and non-shifted normal distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

As we know, Normal distribution support goes from $-\infty$ to $+\infty$. What happens considerig a distribution with a limited support?   
Let us consider the Uniform distribution.

First of all we have to define its probability functions:

```{r qtdshift_unif}
sdunif <- dshift("unif")
spunif <- pshift("unif")
squnif <- qshift("unif")
srunif <- rshift("unif")
```

Now the newly created functions are ready for computing from shifted probability models.  
We limited Uniform support between 6 and 12.

* Density function, `sdunif()`:

    ```{r sdunif_example}
x <- seq(4, 14, len = 100)
not_shifted <- sdunif(x = x, min = 6, max = 12)
positive_shifted <- sdunif(x = x, min = 6, max = 12, shift = 1)
negative_shifted <- sdunif(x = x, min = 6, max = 12, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r sdunif_plot}
df <- data.frame(x=x, 
                 sdunif_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sdunif_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Density plot for shifted and non-shifted uniform distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Probability function, `spunif()`:

    ```{r spunif_example}
not_shifted <- spunif(q = x, min = 6, max = 12)
positive_shifted <- spunif(q = x, min = 6, max = 12, shift = 1)
negative_shifted <- spunif(q = x, min = 6, max = 12, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r spunif_plot}
df <- data.frame(x=x, 
                 spunif_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=spunif_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Quantile") + ylab("Probability") +
  ggtitle("Probability plot for shifted and non-shifted uniform distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Quantile function, `squnif()`:

    ```{r squnif_example}
    x <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
not_shifted <- squnif(p = x, min = 6, max = 12)
positive_shifted <- squnif(p = x, min = 6, max = 12, shift = 1)
negative_shifted <- squnif(p = x, min = 6, max = 12, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r squnif_plot}
df <- data.frame(x=x, 
                 squnif_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=squnif_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3",  
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  ylab("Quantile") + xlab("Probability") +
  ggtitle("Quantile plot for shifted and non-shifted uniform distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Random generator function, `srunif()`:

    ```{r srunif_example}
    x <- 10000
not_shifted <- srunif(n = x, min = 6, max = 12, set_seed = T)
positive_shifted <- srunif(n = x, min = 6, max = 12, shift = 3, set_seed = T)
negative_shifted <- srunif(n = x, min = 6, max = 12, shift = -3, set_seed = T)
    ```
    We can visualize these results by plotting them:

    ```{r srunif_plot}
df <- data.frame(srunif_type = factor(c(rep("not_shifted", times=x),
                    rep("positive_shifted", times=x),
                    rep("negative_shifted", times=x)),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- pl <- ggplot(data = df, mapping = aes(x=value, group=srunif_type, fill=srunif_type)) +
  geom_density(alpha=0.6) +
    scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Density plot for shifted and non-shifted uniform distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)

    ```

As we see, Uniform support changes according to the transformation (the shift) applied to the distribution.

## Extending the computation

In his article, Andrea Spanò use the newly defined probability distribution functions as a base for defining a function for maximum likelihood estimate for the truncated normal distribution. I demonstrate that it is possible also for shifted normal distribution.
Let us see how.

We start from the previously defined fucntion `sdnorm()`, that will be used within an estimator function:

```{r max_likelihood_est}
ltnorm <- function(x, shift = 0) {
  theta <- c(mean(x), sd(x))
  ll <- function(theta, x, shift = 0) {
    mean <- theta[1]
    sd <- theta[2]
    ld <- sdnorm(x = x, mean = mean, sd = sd, shift=shift, log=TRUE)
    -sum(ld)
    } 
  optim(par = theta, fn = ll, x = x, shift=shift)[["par"]]
}
```

```{r max_likelihood_est_example}
# generate a random sample from a shifted normal distribution
x <- srnorm(n = 1000, mean = 5, sd = 2, shift = 1)
# maximum likelihood estimate for the shifted normal distribution
ltnorm(x = x, shift  = 1)
```

## Bibliography

Pace L., Salvan A. . _Introduzione alla Statistica_. CEDAM. 2010

Spano S. .  _R package qdist: A functional programming approach to truncated probability ditributions_. url

Wickham H. . _Advanced r_. http://adv-r.had.co.nz/. Accessed: 2016-08-24.